---
title: "ROH/Heterozygosity Hypotheses"
author: "Anna Brueniche-Olsen and Ken Kellner"
date: "July 14, 2017"
output: pdf_document
---

#Hypothesis 1

Threatened species have lower genome-wide genetic diversity ($\theta$), a larger number of ROHs, a larger mean ROH length, and a larger proportion of their genome in ROHs than non-threatened species, 

```{r echo=FALSE,fig.height=6}
roh20 <- read.csv('roh20X_latlong.csv',header=T)[1:63,]

het <- log(1000*roh20$heterozygosity..H_genome.)
redlist <- roh20$Red_list_groups

keep <- redlist!='D'

het.sub <- het[keep]
redlist.sub <- redlist[keep]

par(mfrow=c(2,2))

mns <- c(mean(het.sub[redlist.sub=='T']),mean(het.sub[redlist.sub=='N']))
sds <- c(sd(het.sub[redlist.sub=='T']),sd(het.sub[redlist.sub=='N']))
cnt <- table(redlist.sub)[-1]
se <- sds/sqrt(cnt)

x <- barplot(mns,col=c('red','blue'),names=c('Threatened','Not Threatened'),ylab="log(heterozygosity*1000)±SE",
        main = "Heterozygosity",ylim=c(0,1.1))

x <- as.vector(x)

segments(x,mns-se,x,mns+se)

out <- t.test(het.sub~redlist.sub,alternative="greater")
text(x[1],0.8,paste('p =',round(out$p.value,3)),cex=1.3)

text(x,mns+se+0.03,cnt)

numrohs <- log(roh20$number.of.ROHs)[keep]
lenrohs <- log(roh20$mean.ROH.length..Kbp.)[keep]
proprohs <- log(1000*roh20$total.number.of.SNPs.in.ROHs/roh20$total.number.of.SNPs)[keep]

#Number of ROHs
mns <- c(mean(numrohs[redlist.sub=='T']),mean(numrohs[redlist.sub=='N']))
sds <- c(sd(numrohs[redlist.sub=='T']),sd(numrohs[redlist.sub=='N']))
cnt <- table(redlist.sub)[-1]
se <- sds/sqrt(cnt)

x <- barplot(mns,col=c('red','blue'),names=c('Threatened','Not Threatened'),ylab="log(number of ROHs)±SE",
             main = "Number of ROHs",ylim=c(0,6))
x <- as.vector(x)

segments(x,mns-se,x,mns+se)

out <- t.test(numrohs~redlist.sub,alternative="less")
text(x[2],5.5,paste('p =',round(out$p.value,3)),cex=1.3)
text(x,mns+se+0.1,cnt)

#Mean ROH length
mns <- c(mean(lenrohs[redlist.sub=='T']),mean(lenrohs[redlist.sub=='N']))
sds <- c(sd(lenrohs[redlist.sub=='T']),sd(lenrohs[redlist.sub=='N']))
cnt <- table(redlist.sub)[-1]
se <- sds/sqrt(cnt)

x <- barplot(mns,col=c('red','blue'),names=c('Threatened','Not Threatened'),ylab="log(mean ROH length)±SE",
             main = "Mean ROH Length",ylim=c(0,5))
x <- as.vector(x)

segments(x,mns-se,x,mns+se)

out <- t.test(lenrohs~redlist.sub,alternative="less")
text(x[2],4.5,paste('p =',round(out$p.value,3)),cex=1.3)
text(x,mns+se+0.1,cnt)

#Prop ROHs (based on SNPs)
mns <- c(mean(proprohs[redlist.sub=='T']),mean(proprohs[redlist.sub=='N']))
sds <- c(sd(proprohs[redlist.sub=='T']),sd(proprohs[redlist.sub=='N']))
cnt <- table(redlist.sub)[-1]
se <- sds/sqrt(cnt)

x <- barplot(mns,col=c('red','blue'),names=c('Threatened','Not Threatened'),ylab="log(1000*proportion ROHs)±SE",
             main = "Proportion in ROHs",ylim=c(-0.3,3))
x <- as.vector(x)

segments(x,mns-se,x,mns+se)

out <- t.test(proprohs~redlist.sub,alternative="less")
text(x[2],1.5,paste('p =',round(out$p.value,3)),cex=1.3)
text(x,mns+se+0.1,cnt)

```

\pagebreak

#Hypothesis 2

Genetic diversity increases and the number and size of ROHs decreases with trophic level from carnivore, to herbivore, to omnivore.

```{r echo=FALSE, fig.height=6}

par(mfrow=c(2,2))

het <- log(10000*roh20$heterozygosity..H_genome.)

diet <- roh20$Class.of.Consumption
tab <- data.frame(het=het,diet=diet)

mns <- with(tab,tapply(het,diet,mean))
sds <- with(tab,tapply(het,diet,sd))
cnt <- table(diet)
se <- sds/sqrt(cnt)

x <- barplot(mns,names=c('Carnivore','Herbivore','Omnivore'),
             col=c('Red','Green','Blue'),
             ylab="log(10000*heterozygosity)±SE",
             ylim=c(0,4),main='Heterozygosity')

segments(x,mns+se,x,mns-se)

test = aov(het~diet)
text(x[1],3.5,paste('p =',round(summary(test)[[1]][1,5],3)),cex=1.3)
text(x,mns+se+0.1,cnt)

numrohs <- log(roh20$number.of.ROHs)
lenrohs <- log(roh20$mean.ROH.length..Kbp.)
proprohs <- log(1000*roh20$total.number.of.SNPs.in.ROHs/roh20$total.number.of.SNPs)

tab <- data.frame(numrohs=numrohs,diet=diet)
mns <- with(tab,tapply(numrohs,diet,mean))
sds <- with(tab,tapply(numrohs,diet,sd))
cnt <- table(diet)
se <- sds/sqrt(cnt)

x <- barplot(mns,names=c('Carnivore','Herbivore','Omnivore'),
             col=c('Red','Green','Blue'),
             ylab="log(number of ROHs)±SE",
             ylim=c(0,6.5),main='Number of ROHs')
segments(x,mns+se,x,mns-se)

test = aov(numrohs~diet)

text(x[1]+0.1,6.2,paste('p =',round(summary(test)[[1]][1,5],3)),cex=1.3)
text(x,mns+se+0.2,cnt)
##
tab <- data.frame(lenrohs=lenrohs,diet=diet)
mns <- with(tab,tapply(lenrohs,diet,mean))
sds <- with(tab,tapply(lenrohs,diet,sd))
cnt <- table(diet)
se <- sds/sqrt(cnt)

x <- barplot(mns,names=c('Carnivore','Herbivore','Omnivore'),
             col=c('Red','Green','Blue'),
             ylab="log(length of ROHs)±SE",
             ylim=c(0,6),main='Mean Length of ROHs')
segments(x,mns+se,x,mns-se)

test = aov(lenrohs~diet)

text(x[1]+0.1,5.7,paste('p =',round(summary(test)[[1]][1,5],3)),cex=1.3)
text(x,mns+se+0.1,cnt)
##
tab <- data.frame(proprohs=proprohs,diet=diet)
mns <- with(tab,tapply(proprohs,diet,mean))
sds <- with(tab,tapply(proprohs,diet,sd))
cnt <- table(diet)
se <- sds/sqrt(cnt)

x <- barplot(mns,names=c('Carnivore','Herbivore','Omnivore'),
             col=c('Red','Green','Blue'),
             ylab="log(1000*proportion ROHs)±SE",
             ylim=c(0,2),main='Proportion of ROHs')
segments(x,mns+se,x,mns-se)

test = aov(proprohs~diet)

text(x[1]+0.1,1.9,paste('p =',round(summary(test)[[1]][1,5],3)),cex=1.3)
text(x,mns+se+0.1,cnt)
```

\pagebreak

#Hypothesis 3

Genetic diversity increases with decreasing latitude so that species near the poles have lower genetic diversity than species closer to equator

```{r echo=FALSE,fig.height=6}
par(mfrow=c(2,2))

lat <- roh20$lat

het <- log(10000*roh20$heterozygosity..H_genome.)
numrohs <- log(roh20$number.of.ROHs)
lenrohs <- log(roh20$mean.ROH.length..Kbp.)
proprohs <- log(1000*roh20$total.number.of.SNPs.in.ROHs/roh20$total.number.of.SNPs)

plot(abs(lat),het,pch=19,
             ylab="log(10000*heterozygosity)",
             xlab="abs(latitude) of distribution centroid"
             ,main='Heterozygosity')

test = lm(het~abs(lat))
abline(test)
text(0,4.5,paste('p =',round(summary(test)[[4]][2,4],3)),cex=1.3,adj=c(0,0))

##
plot(abs(lat),numrohs,pch=19,
             ylab="log(number of ROHs)",
             xlab="abs(latitude) of distribution centroid"
             ,main='Number of ROHs')

test = lm(numrohs~abs(lat))
abline(test)
text(50,10,paste('p =',round(summary(test)[[4]][2,4],3)),cex=1.3,adj=c(0,0))

##
plot(abs(lat),lenrohs,pch=19,
             ylab="log(length of ROHs)",
             xlab="abs(latitude) of distribution centroid"
             ,main='Mean Length of ROHs')

test = lm(lenrohs~abs(lat))
abline(test)
text(0,7.1,paste('p =',round(summary(test)[[4]][2,4],3)),cex=1.3,adj=c(0,0))

##
plot(abs(lat),proprohs,pch=19,
             ylab="log(1000*proportion ROHs)",
             xlab="abs(latitude) of distribution centroid"
             ,main='Proportion of ROHs')

test = lm(proprohs~abs(lat))
abline(test)
text(50,-6,paste('p =',round(summary(test)[[4]][2,4],3)),cex=1.3,adj=c(0,0))
```

\pagebreak

#Hypothesis 4

Effects are stronger when factors are combined; i.e., if a species is carnivore and threatened it will have a lower genetic diversity, more and longer ROHs than threatened herbivores. 

```{r echo=FALSE,fig.height=6}

par(mfrow=c(2,2))

keep <- redlist=='T'&diet%in%c('C','H')

het <- log(10000*roh20$heterozygosity..H_genome.)[keep]
numrohs <- log(roh20$number.of.ROHs)[keep]
lenrohs <- log(roh20$mean.ROH.length..Kbp.)[keep]
proprohs <- log(1000*roh20$total.number.of.SNPs.in.ROHs/roh20$total.number.of.SNPs)[keep]
diet.sub <- diet[keep]


##
tab <- data.frame(het=het,diet=diet.sub)
mns <- with(tab,tapply(het,diet,mean))[c(-3)]
sds <- with(tab,tapply(het,diet,sd))[c(-3)]
cnt <- table(diet.sub)[c(-3)]
se <- sds/sqrt(cnt)

x <- barplot(mns,names=c('Carnivore','Herbivore'),
             col=c('Red','Green'),
             ylab="log(10000*heterozygosity)±SE",
             ylim=c(0,4),main='Heterozygosity')
segments(x,mns+se,x,mns-se)

text(x,mns+se+0.3,cnt)

test = t.test(het~diet.sub,alternative='less')
text(x[1],3.5,paste('p =',round(test$p.value,3)),cex=1.3)

##
tab <- data.frame(numrohs=numrohs,diet=diet.sub)
mns <- with(tab,tapply(numrohs,diet,mean))[c(-3)]
sds <- with(tab,tapply(numrohs,diet,sd))[c(-3)]
cnt <- table(diet.sub)[c(-3)]
se <- sds/sqrt(cnt)

x <- barplot(mns,names=c('Carnivore','Herbivore'),
             col=c('Red','Green'),
             ylab="log(number of ROHs)±SE",
             ylim=c(0,7),main='Number of ROHs')
segments(x,mns+se,x,mns-se)

test = t.test(numrohs~diet.sub,alternative='greater')
text(x[1]+0.7,6.5,paste('p =',round(test$p.value,3)),cex=1.3)


text(x,mns+se+0.3,cnt)
##
tab <- data.frame(lenrohs=lenrohs,diet=diet.sub)
mns <- with(tab,tapply(lenrohs,diet,mean))[c(-3)]
sds <- with(tab,tapply(lenrohs,diet,sd))[c(-3)]
cnt <- table(diet.sub)[c(-3)]
se <- sds/sqrt(cnt)

x <- barplot(mns,names=c('Carnivore','Herbivore'),
             col=c('Red','Green'),
             ylab="log(mean length of ROHs)±SE",
             ylim=c(0,6),main='Mean Length of ROHs')
segments(x,mns+se,x,mns-se)

test = t.test(lenrohs~diet.sub,alternative='greater')
text(x[1],5.5,paste('p =',round(test$p.value,3)),cex=1.3)

text(x,mns+se+0.3,cnt)

##
tab <- data.frame(proprohs=proprohs,diet=diet.sub)
mns <- with(tab,tapply(proprohs,diet,mean))[c(-3)]
sds <- with(tab,tapply(proprohs,diet,sd))[c(-3)]
cnt <- table(diet.sub)[c(-3)]
se <- sds/sqrt(cnt)

x <- barplot(mns,names=c('Carnivore','Herbivore'),
             col=c('Red','Green'),
             ylab="log(1000*proportion of ROHs)±SE",
             ylim=c(0,4),main='Proportion of ROHs')
segments(x,mns+se,x,mns-se)

test = t.test(proprohs~diet.sub,alternative='greater')
text(x[1]+0.7,3.5,paste('p =',round(test$p.value,3)),cex=1.3)

text(x,mns+se+0.3,cnt)


```
