---
title: "ROH Analysis 10x"
author: "Anna Brueniche-Olsen and Ken Kellner"
output: pdf_document
---

```{r,echo=F}

setwd('8_R_analyses')
source('gen_filepath.R')

roh20 <- read.csv(datapath('roh10X_full.csv'),header=T)
roh20$mean.ROH.length..Kbp.[roh20$mean.ROH.length..Kbp.=='#DIV/0!'] <- NA

#Generate response variables (log-transform)
het <- log(roh20$hetero)

numrohs <- log(roh20$ROHnumber+1)
#numrohs[which(is.infinite(numrohs))] <- NA

lenrohs <- log(as.numeric(as.character(roh20$mean.ROH.length..Kbp.))+1)
#lenrohs[which(is.infinite(lenrohs))] <- NA

proprohs <- log((roh20$SNVROHnumber+1)/(roh20$sites+1))
#proprohs[which(is.infinite(proprohs))] <- NA


#Test normality
# shapiro.test(het)
# shapiro.test(numrohs) #Fails
# shapiro.test(lenrohs)
# shapiro.test(proprohs) #Fails

#Clean up covariates
redlist <- relevel(roh20$Red_list_groups,ref='N')
diet <- relevel(roh20$Class.of.Consumption,ref='O')
lat <- scale(abs(roh20$lat))
mass <- scale(log(roh20$mass))

cutoff1 <- 1000
cutoff2 <- 100000
masscat <- rep(NA,length(mass))
masscat[roh20$mass<cutoff1] <- 'S'
masscat[roh20$mass>=cutoff1&roh20$mass<cutoff2] <- 'M'
masscat[roh20$mass>=cutoff2] <- 'L'
masscat <- as.factor(masscat)
masscat <- relevel(masscat,ref='S')

#Build complete dataset
data <- data.frame(het=het,numrohs=numrohs,lenrohs=lenrohs,proprohs=proprohs,
                   redlist=redlist,diet=diet,lat=lat,mass=mass,masscat=masscat)

####################
```

# Heterozygosity

Multiple regression output:

```{r,echo=F}
data.sub <- data[data$redlist!='D',] #Remove species w/out data
data.sub$redlist <- factor(data.sub$redlist)

test = lm(het~lat+mass+diet*redlist,data=data.sub)
#summary(test)

knitr::kable(summary(test)$coefficients,digits=3)
```

# Mean ROH Length

Multiple regression output:

```{r,echo=F}
data.sub <- data[data$redlist!='D',] #Remove species w/out data
data.sub$redlist <- factor(data.sub$redlist)

test = lm(lenrohs~lat+mass+diet*redlist,data=data.sub)

knitr::kable(summary(test)$coefficients,digits=3)
```


\pagebreak


# Number of ROHs

Kruskal-Wallis test for threatened status effect:

```{r,echo=F}
data.sub <- data[data$redlist!='D',] #Remove species w/out data
data.sub$redlist <- factor(data.sub$redlist)
kruskal.test(data.sub$numrohs,data.sub$redlist)
````

Kruskal-Wallis test for diet class effect:

```{r,echo=F}
kruskal.test(data$numrohs,data$diet)
````

Spearman's rank correlation test for body mass effect:

```{r,echo=F}
cor.test(data$numrohs,data$mass,method='spearman')
````

Spearman's rank correlation test for latitude effect:

```{r,echo=F}
cor.test(data$numrohs,data$lat,method='spearman')
````

# Proportion of SNVs in ROHs

Kruskal-Wallis test for threatened status effect:

```{r,echo=F}
data.sub <- data[data$redlist!='D',] #Remove species w/out data
data.sub$redlist <- factor(data.sub$redlist)
kruskal.test(data.sub$proprohs,data.sub$redlist)
````

Kruskal-Wallis test for diet class effect:

```{r,echo=F}
kruskal.test(data$proprohs,data$diet)
````

Spearman's rank correlation test for body mass effect:

```{r,echo=F}
cor.test(data$proprohs,data$mass,method='spearman')
````

Spearman's rank correlation test for latitude effect:

```{r,echo=F}
cor.test(data$proprohs,data$lat,method='spearman')
````
